---
title: "Part 2: Calculating Comorbidities"
author: "Campbell/Langasek"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summarizing Comorbidities (C)

## Due Date: August 24 (Monday of Week 9)

## Purpose 

1) To summarize chronic illnesses (choose 2) for the LACE Comorbidity (C) score
2) Bring each element (L, A, C, E) together into a LACE score for each patient
3) Assess how well the LACE score predicts 30-day admission risk using simple analysis.

##Background 

LACE has several comorbidities that it summarizes to help predict risk.  It turns out that they do not provide specific code sets to match these to the primary set, and that there may be other comorbidities that are important for this work.  Prior analysis has found that in your patient population Dementia, COPD, CAD, Stroke, Diabetes, Complicated Diabetes, Cancer, Hepatitis C, and Asthma are the most likely candidates to be associated with readmission risk.   

However, diagnoses are coded as very specific ICD-9 codes (patient_diagnosis.icd9code); each grouped diagnosis represents many of these codes.  

For your task, you will need to find acceptable code groupings for 2 of the following diagnoses and summarize these for each patient admission.

- CAD (414)
- CHF (428)
- Asthma (493)
- Hepatitis C (070,V02.62,V12.09)
- Dementia (290)
- Stroke(431,433-5)
- Depression (296,300.4,309.0,309.28311)
- Peripheral Vascular Disease (440.2,443)

## Deliverables

1) Summarized comorbidities and total LACE scores for each patient as a table.
2) Source code for your summarization and analysis.

## Suggested Steps

1) Find the ICD9 code groupings you will use for your two comorbidities. We will discuss some in class.

2) For your two selected comorbidities, check whether a patient has that diagnosis (from the `patient_diagnosis` table) and store the outcome in your analytic file as a binary variable (such as 1 = patient has comorbidity and 0 = patient does not have comorbidity). *Hint*: since each patient may have multiple diagnoses, you will need to find a way to summarize/collapse associated diagnoses into a single binary variable for each comorbidity.

3) Calculate the total comorbidity score for each patient as per the LACE appendix. *Optional*: If you think you'd like to weight some comorbidities over others, you can adjust the weights/values for those comorbidities.


```{r}
#load the RSQLite library
library(RSQLite)
library(here)
library(dplyr)
```

```{r}

#connect to our database
con <- dbConnect(drv=SQLite(), dbname=here("data/patient1.sqlite"))

dbListTables(con)
```

```{r}
#Use union statements to combine queries selecting relevent ICD9 codes and then splitting the case statement to account for points and diagnosis name

sqlStatement <- "SELECT patientid, ICD9Code, substr(DxGroup,1,2) as DxGroupCode, substr(DxGroup,4,1) as DxPoints, substr(DxGroup,6) as DxGroup
                FROM 
                (SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '02_2_Congestive Heart Failure' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('398.91','402.01','402.11','402.91','404.01','404.03','404.11','404.13','404.91','404.93','425.4','425.5','425.6','425.7','425.8','425.9','428') OR ICD9Code LIKE '428.%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '01_1_Myocardial Infarction' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('410','412') OR ICD9Code LIKE '410.%' OR ICD9Code LIKE '412.%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '03_1_Peripheral Vascular Disease' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('093.0','437.3','440','441','443.1','443.2','443.3','443.4','443.5','443.6','443.7','443.8','443.9','447.1','557.1','557.9','V43.4') OR ICD9Code LIKE '440.%' OR ICD9Code LIKE '441.%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '04_1_Cerebrovascular Disease' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('362.34') OR ICD9Code LIKE '430%' OR ICD9Code LIKE '431%' OR ICD9Code LIKE '432%' OR ICD9Code LIKE '433%' OR ICD9Code LIKE '434%' OR ICD9Code LIKE '435%' OR ICD9Code LIKE '436%' OR ICD9Code LIKE '437%' OR ICD9Code LIKE '438%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '05_3_Dementia' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('290','294.1','331.2') OR ICD9Code LIKE '290.%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '06_2_Chronic Pulmonary Disease' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('416.8','416.9','506.4','508.1','508.8') OR ICD9Code LIKE '490%' OR ICD9Code LIKE '491%' OR ICD9Code LIKE '492%' OR ICD9Code LIKE '493%' OR ICD9Code LIKE '494%' OR ICD9Code LIKE '495%' OR ICD9Code LIKE '496%' OR ICD9Code LIKE '497%' OR ICD9Code LIKE '498%' OR ICD9Code LIKE '499%' OR ICD9Code LIKE '500%' OR ICD9Code LIKE '501%' OR ICD9Code LIKE '502%' OR ICD9Code LIKE '503%' OR ICD9Code LIKE '504%' OR ICD9Code LIKE '505%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '07_3_Rheumatic Disease' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('446.5','710.0','710.1','710.2','710.3','710.4','714.0','714.1','714.2','714.8','725') OR ICD9Code LIKE '725.%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '08_0_Peptic Ulcer Disease' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code LIKE '531%' OR ICD9Code LIKE '532%' OR ICD9Code LIKE '533%' OR ICD9Code LIKE '534%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '09_2_Mild Liver Disease' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('070.22', '070.23', '070.32', '070.33', '070.44', '070.54', '070.6', '070.9','573.3', '573.4', '573.8', '573.9', 'V42.7') OR ICD9Code LIKE '570%' OR ICD9Code LIKE '571%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '10_1_Diabetes Without Chronic Complication' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('250.0','250.1','250.2','250.3','250.8','250.9')
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '11_2_Diabetes With Chronic Complication' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('250.4','250.5','250.6','250.7')
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '12_0_Hemiplegia or paraplegia' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('334.1','344.0','344.1','344.2','344.3','344.4','344.5','344.6','344.9') OR ICD9Code LIKE '342%' OR ICD9Code LIKE '343%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '13_0_Renal Disease' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('403.01', '403.11', '403.91', '404.02','404.03', '404.12', '404.13', '404.92','404.93','583.0','583.1','583.2','583.3','583.4','583.5','583.6','583.7','588.0','V42.0','V45.1') OR ICD9Code LIKE '582%' OR ICD9Code LIKE '412.%' OR ICD9Code LIKE '585%' OR ICD9Code LIKE '586%' OR ICD9Code LIKE 'V56%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '14_2_Cancer' END as DxGroup
                 FROM patient_diagnosis
                  WHERE (ICD9Code = '238.6' OR ICD9Code LIKE '14_%' OR ICD9Code LIKE '15_%' OR ICD9Code LIKE '16_%' OR ICD9Code LIKE '17_%' OR ICD9Code LIKE '18_%' OR ICD9Code LIKE '190%' OR ICD9Code LIKE '191%' OR ICD9Code LIKE '192%' OR ICD9Code LIKE '193%' OR ICD9Code LIKE '194%' OR ICD9Code LIKE '195' OR ICD9Code LIKE '20_%') AND (ICD9Code NOT LIKE '173%' OR ICD9Code NOT LIKE '195.9%' OR ICD9Code NOT LIKE '209%')
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '15_4_Moderate or severe liver disease' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code IN ('456.0','456.1','456.2','572.2','572.3','572.4','572.5','572.6','572.7','572.8') 
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '16_6_Metastatic Carcinoma' END as DxGroup
                 FROM patient_diagnosis
                 WHERE ICD9Code LIKE '196%' OR ICD9Code LIKE '197%' OR ICD9Code LIKE '198%' OR ICD9Code LIKE '199%'
                  UNION
                  SELECT patientid, ICD9Code, Case WHEN ICD9Code IS NOT NULL THEN '17_4_AIDS/HIV' END as DxGroup
                 FROM patient_diagnosis
                  WHERE ICD9Code LIKE '042%' OR ICD9Code LIKE '043%' OR ICD9Code LIKE '044%')" 

queryScoreResult <- tbl(con, sql(sqlStatement))

queryScoreResult
```

```{r}
tableResultScore <-collect(queryScoreResult)

#Save the results in a table
dbWriteTable(conn = con, 
             name="comorbidityScoreTable",
             value = tableResultScore, 
             overwrite=TRUE)

sqlStatement <- "SELECT * FROM comorbidityScoreTable"

queryResult <- tbl(con, sql(sqlStatement))

queryResult
```
```{r}
#check the table results
sqlStatement <- "SELECT patientID, SUM(DxPoints) 
                FROM comorbidityScoreTable
                GROUP BY patientid"

queryResult <- tbl(con, sql(sqlStatement))

queryResult
```

```{r}
#Combine points to form LACE score
sqlStatement <- "SELECT patientid, (Length_of_stay+Admission_acuity+CASE WHEN ifnull(coPts,0)<3 THEN ifnull(coPts,0) ELSE 5 END+ED_visits_score) LACE FROM (
                  (SELECT * 
                  FROM analyticTable 
                  WHERE index_admit=1) LEFT JOIN 
                  (SELECT patientID, SUM(DxPoints) coPts
                  FROM comorbidityScoreTable
                  GROUP BY patientid) USING (patientID)
                )"

queryResult <- tbl(con, sql(sqlStatement))

queryResult
```

```{r}
dbDisconnect(con)
```
