---
title: 'Problem Set 3: Modeling'
author: "Campbell/Langasek"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Due Date: August 31 (Monday of Week 10)

1) Calculate the LACE scores for each patient. 

2 Run a basic analysis of the LACE variables to predict readmission risk using your analytic file. Some kinds of analysis you can do: contingency tables ([more info here](http://www.statmethods.net/stats/frequencies.html) ) or logistic regression (More info [here](https://www.r-bloggers.com/evaluating-logistic-regression-models/) and [here](https://www.analyticsvidhya.com/blog/2015/11/beginners-guide-on-logistic-regression-in-r/) ). How well does the LACE score predict readmission risk? Can you compare the predictive power of each variable (L, A, C, E) to predict risk? Begin to think about what visualizations and analysis supports your conclusions.

3) *Optional*: Try changing the ‘weights’ of each of the individual scores to calculate a different LACE score and see the difference in predictive validity.

```{r}
#load the RSQLite library
library(RSQLite)
library(here)
library(dplyr)
library(tidyverse)
library(broom)
library(caret)
library(ggplot2)
library(plotly)
library(ROCR)
```

```{r}

#connect to our database
con <- dbConnect(drv=SQLite(), dbname=here("data/patient1.sqlite"))

dbListTables(con)
```

```{r}
#Combine points to form LACE score
sqlStatement <- "SELECT patientid, Length_of_stay, Admission_acuity, ifnull(coPts,0) comorbidity, CASE WHEN ifnull(coPts,0)<3 THEN ifnull(coPts,0) ELSE 5 END coPts, ED_visits_score, (Length_of_stay+Admission_acuity+CASE WHEN ifnull(coPts,0)<3 THEN ifnull(coPts,0) ELSE 5 END+ED_visits_score) LACE, readmit30 FROM (
                  (SELECT * 
                  FROM analyticTable 
                  WHERE index_admit=1) LEFT JOIN 
                  (SELECT patientID, SUM(DxPoints) coPts
                  FROM comorbidityScoreTable
                  GROUP BY patientid) USING (patientID)
                )"


queryResult <- tbl(con, sql(sqlStatement))

queryResult

tableResult <-collect(queryResult)

saveRDS(tableResult, file='data/LACE.RDS')
```




```{r}
set.seed(111)

LACE_pt <- readRDS("data/LACE.RDS")

LACE_pt$readmit30 <-factor(LACE_pt$readmit30,labels=c('no','yes'))
summary(LACE_pt)
```
```{r}
#grab indices of the dataset that represent 80% of the data
trainingIndices <- createDataPartition(y = LACE_pt$readmit30, p=.80,
                                       list=FALSE)

#show the first few training indices
trainingIndices[1:10]

#select the rows
trainData <- LACE_pt[trainingIndices,]

#confirm the number of rows (should be 80)
nrow(trainData)

#build our test set using the R-indexing
#using the "-" operator
testData <- LACE_pt[-trainingIndices,]

#confirm the number of rows 
nrow(testData)
```
```{r}
#show variable names in analytic data.frame
colnames(trainData)

#run a simple logistic regression model just using age and gender
#we can cast gender as categorical data using factor()

LACEModel <- glm(readmit30 ~ Length_of_stay + Admission_acuity + coPts + ED_visits_score, data= trainData, family="binomial")
summary(LACEModel)
```
```{r}
#Summarize the model
tidy(LACEModel)
```
```{r}
#grab coefficients themselves
coef(LACEModel)
```
```{r}
modelPredictedProbabilities <- predict(LACEModel, newdata=testData, type = "response")

##add the modelPredictedProbabilities as a column in testData
testData <- data.frame(testData, predProb=modelPredictedProbabilities)
#testDataAugment <- augment(ageGenderModel)

testData %>% ggplot(aes(x=LACE, y=predProb, color=Admission_acuity)) + geom_point()
```
```{r}
#predict the logit instead for our testData
modelPredictedLogOddsRatio <- predict(LACEModel,newdata = testData)

#add as another column in our table
testDataPred <- data.frame(testData, predLogit = modelPredictedLogOddsRatio)

#plot the age versus logit (coloring by gender)
testDataPred %>% ggplot(aes(x=LACE, y=predLogit, color=Admission_acuity)) + geom_point()
```
```{r}
#transform the logit to the predictedOdds ratio
modelPredictedOddsRatio <- exp(modelPredictedLogOddsRatio)
modelPredictedOddsRatio[1:10]

#add as column in our table
testDataPred <- data.frame(testDataPred, predOR = modelPredictedOddsRatio)

#plot Odds ratio versus age
testDataPred %>% ggplot(aes(x=LACE, y=predOR, col=Admission_acuity)) + geom_point() + xlim(0,25) + ylim(0,0.5)
```
```{r}
exp(coef(LACEModel))
```

```{r}
modelPredictedProbabilities %>% data.frame() %>% ggplot(aes(x=modelPredictedProbabilities)) + geom_histogram()
```
```{r}
modelPredictions <- ifelse(modelPredictedProbabilities < 0.13, "N", "Y")
modelPredictions[1:10]
```
```{r}
truthPredict <- table(testData$readmit30, modelPredictions)
truthPredict
```
```{r}
totalCases <- sum(truthPredict)
misclassified <- truthPredict[1,2] + truthPredict[2,1]
misclassified
accuracy <- (totalCases - misclassified) / totalCases
accuracy
```
```{r}

confusionMatrix(factor(testData$readmit30), factor(modelPredictions), positive = "Y")
```
```{r}
truth <- factor(testData$readmit30, levels=c("Y","N"))
#seq() function allows us to make a sequence of cutoffs
cutoffs <- seq(from=0.01,to=1, by = 0.01)
cutoffs
```

```{r}
# this function will produce a confusion matrix
# for a particular cutoff
run_cutoff <- function(cutoff){
  #Do thresholding with cutoff       
  predictions <- ifelse(modelPredictedProbabilities < cutoff, "N", "Y")
  #run confusion matrix on predictions
  caret::confusionMatrix(
          factor(predictions, levels=c("Y", "N")), 
                  truth, 
                  positive="Y")
  }
```

```{r}
#run our function on all of our cutoffs
#the output is a list of confusion matrices
confusion_matrices <- lapply(cutoffs, run_cutoff)

#get the sensitivity out of the list of confusion matrices
sensitivities <- lapply(confusion_matrices, function(x){return(x$byClass["Sensitivity"])})

#get the sensitivity out of the list of confusion matrices
specificities <- lapply(confusion_matrices, function(x){return(x$byClass["Specificity"])})

#make sensitivities and specificities into vectors
sensitivities <- unlist(sensitivities)
specificities <- unlist(specificities)

#produce a data frame with all of the cutoffs
cutoff_frame <- data.frame(cutoff=cutoffs, sensitivity =sensitivities, specificity = specificities)
cutoff_frame
```


```{r}
dbDisconnect(con)
```